An Open-Source Implementation of the CMPS Algorithm for Assessing Similarity of Bullets

### 2022-03-29 submissions/resubmission-ju-hofmann2/ju-hofmann.Rmd
update variable names in ju-hofmann.R

CMPS_hamby252_results_seg <- all_results_container$h252_seg
CMPS_hamby252_results_npeak <- all_results_container$h252_npeak
CMPS_hamby44_results_seg <- all_results_container$h44_seg
CMPS_hamby44_results_npeak <- all_results_container$h44_npeak
rf.results <- all_results_container$rf

@@

```{r rf-data-process, echo=FALSE}
rf.tt <- lapply(rf.results, function(rf_result) {
  with(rf_result, {
    tibble(rf.diff, rf.max, rf.maxbar) %>%
      apply(2, compute_ss_ratio, label = type_truth, MS = FALSE)
  })
}) %>%
  do.call(rbind, .) %>%
  as_tibble()

colnames(rf.tt) <- c("cmps.diff", "cmps.max", "cmps.maxbar")

rf.tt$cmps.diff_scaled <- rf.tt$cmps.diff

rf.tt$study <- c("Hamby 252", "Hamby 44")
rf.tt <- rf.tt %>% pivot_longer(
  cols = c("cmps.diff", "cmps.max", "cmps.maxbar", "cmps.diff_scaled"),
  names_to = "Metric"
)

rf.tt$npeak <- "RF score* (Hare et al)"
```

```{r ss-original-cmps, echo=FALSE}
result.container <- list()

result.container$npeak <-
  c("5* (Chen et al)", "5-3-1* (Chen et al)", "5* (Chen et al)", "5-3-1* (Chen et al)")
result.container$Metric <- c("cmps.max", "cmps.max", "cmps.maxbar", "cmps.maxbar")
result.container$study <- "Hamby 252"

dd.list <- list()

dd.list[[1]] <- tibble(
  score = c(9:13, 16:24),
  freq = c(132, 282, 115, 19, 1, 1, 3, 2, 4, 8, 12, 7, 7, 2),
  type = c(rep("KNM", 5), rep("KM", 9))
)

dd.list[[2]] <- tibble(
  score = c(1:5, 14:23),
  freq = c(7, 343, 168, 24, 7, 1, 1, 2, 3, 6, 7, 8, 8, 7, 3),
  type = c(rep("KNM", 5), rep("KM", 10))
)

dd.list[[3]] <- tibble(
  score = c(6:9, 13:19),
  freq = c(2, 210, 307, 30, 6, 3, 9, 9, 11, 7, 1),
  type = c(rep("KNM", 4), rep("KM", 7))
)

dd.list[[4]] <- tibble(
  score = c(0:2, 8:17),
  freq = c(36, 508, 5, 1, 3, 4, 2, 11, 3, 9, 8, 4, 1),
  type = c(rep("KNM", 3), rep("KM", 10))
)

result.container$dd.list <- dd.list

result.container <- result.container %>%
  as_tibble() %>%
  mutate(
    dd.table = purrr::map(dd.list, function(dd) {
      tibble(
        score = rep(dd$score, dd$freq),
        type_truth = rep(dd$type, dd$freq)
      )
    }),
    value = purrr::map_dbl(dd.table, function(dd) {
      compute_ss_ratio(dd$score, dd$type_truth, MS = FALSE)
    })
  )
```

@@

```{r plot_computation, echo=FALSE, fig.width=8, fig.height = 8, out.width='\\linewidth'}
# 252 - seg
table.252.cmps_seg <- lapply(CMPS_hamby252_results_seg$cmps.table, function(tt) {
  type_truth <- tt$type_truth
  tt %>%
    select(cmps.diff:cmps.maxbar_scaled) %>%
    apply(2, compute_ss_ratio, label = type_truth, MS = FALSE)
}) %>%
  do.call(rbind, .) %>%
  as.data.frame()

table.252.cmps_seg$seg_length <- unlist(CMPS_hamby252_results_seg$seg_length)

table.252.cmps_seg.long <- table.252.cmps_seg %>%
  pivot_longer(
    cols = c(
      cmps.diff, cmps.diff_scaled,
      cmps.max, cmps.maxbar
    ),
    names_to = "Metric"
  )

table.252.cmps_seg.long <- table.252.cmps_seg.long %>%
  mutate(study = "Hamby 252")

@@

p.seg_252_plot <- table.252.cmps_seg.long %>%
  ggplot(aes(x = seg_length + 3 * as.numeric(factor(Metric)) - 9)) +
  geom_segment(aes(
    xend = seg_length + 3 * as.numeric(factor(Metric)) - 9,
    y = value, yend = 0, colour = Metric
  )) +
  geom_point(aes(y = value, color = Metric),
    alpha = 0.8,
    position = position_dodge(width = 0.5), size = 3.5
  ) +
  scale_x_continuous(
    breaks = table.252.cmps_seg$seg_length,
    minor_breaks = NULL
  ) +
  theme_bw() +
  # theme(legend.position="none") +
  ggtitle("Results for Hamby 252") +
  ylab("Sum of Squares Ratio") +
  xlab("Segment Length") +
  scale_colour_manual(values = cols_2, labels = labels) +
  scale_fill_manual(values = cols_1, labels = labels)

# 44 - seg
table.44.cmps_seg <- lapply(CMPS_hamby44_results_seg$cmps.table, function(tt) {
  type_truth <- tt$type_truth
  tt %>%
    select(cmps.diff:cmps.maxbar_scaled) %>%
    apply(2, compute_ss_ratio, label = type_truth, MS = FALSE)
}) %>%
  do.call(rbind, .) %>%
  as.data.frame()

table.44.cmps_seg$seg_length <- unlist(CMPS_hamby44_results_seg$seg_length)

table.44.cmps_seg.long <- table.44.cmps_seg %>%
  pivot_longer(
    cols = c(
      cmps.diff, cmps.diff_scaled,
      cmps.max, cmps.maxbar
    ),
    names_to = "Metric"
  )

table.44.cmps_seg.long <- table.44.cmps_seg.long %>% mutate(
  study = "Hamby 44"
)

table.cmps_seg.long <- bind_rows(
  table.44.cmps_seg.long,
  table.252.cmps_seg.long
)

@@

# npeak - 252
table.252.cmps_npeak <- lapply(CMPS_hamby252_results_npeak$cmps.table, function(tt) {
  type_truth <- tt$type_truth
  tt %>%
    select(cmps.diff:cmps.maxbar_scaled) %>%
    apply(2, compute_ss_ratio, label = type_truth, MS = FALSE)
}) %>%
  do.call(rbind, .) %>%
  as.data.frame()

table.252.cmps_npeak$npeak <- CMPS_hamby252_results_npeak$npeaks_set %>%
  purrr::map(.f = paste, collapse = "-") %>%
  sapply(as.character)

table.252.cmps_npeak.longer <- table.252.cmps_npeak %>%
  as_tibble() %>%
  filter(npeak %in% c("5", "5-3-1", "10-8-6-4-2-1", "10-7-4-2")) %>%
  mutate(
    npeak = factor(
      npeak,
      levels = npeak[order(cmps.diff_scaled, decreasing = FALSE)]
    )
  ) %>%
  pivot_longer(
    cols = c(cmps.diff, cmps.diff_scaled, cmps.max, cmps.maxbar),
    names_to = "Metric"
  )
table.252.cmps_npeak.longer$study <- "Hamby 252"

# npeak - 44
table.44.cmps_npeak <- lapply(CMPS_hamby44_results_npeak$cmps.table, function(tt) {
  type_truth <- tt$type_truth
  tt %>%
    select(cmps.diff:cmps.maxbar_scaled) %>%
    apply(2, compute_ss_ratio, label = type_truth, MS = FALSE)
}) %>%
  do.call(rbind, .) %>%
  as.data.frame()

table.44.cmps_npeak$npeak <- CMPS_hamby44_results_npeak$npeaks_set %>%
  purrr::map(.f = paste, collapse = "-") %>%
  sapply(as.character)

table.44.cmps_npeak.longer <- table.44.cmps_npeak %>%
  as_tibble() %>%
  filter(npeak %in% c("5", "5-3-1", "10-8-6-4-2-1", "10-7-4-2")) %>%
  mutate(
    npeak = factor(
      npeak,
      levels = npeak[order(cmps.diff_scaled, decreasing = FALSE)]
    )
  ) %>%
  pivot_longer(
    cols = c(cmps.diff, cmps.diff_scaled, cmps.max, cmps.maxbar),
    names_to = "Metric"
  )
table.44.cmps_npeak.longer$study <- "Hamby 44"

@@

table.cmps_neapk.long <- bind_rows(
  table.252.cmps_npeak.longer %>% select(all_of(selected_var)),
  table.44.cmps_npeak.longer %>% select(all_of(selected_var)),
  rf.tt %>% select(all_of(selected_var)),
  result.container %>% select(all_of(selected_var))
)

order.level <- table.cmps_neapk.long %>%
  filter(Metric == "cmps.maxbar", study == "Hamby 252") %>%
  {
    .$npeak[order(.$value, decreasing = FALSE)]
  }

table.cmps_neapk.long <- table.cmps_neapk.long %>% mutate(
  npeak = factor(npeak, levels = order.level)
)

npeak_convert <- with(table.cmps_neapk.long, {
  tt <- tibble(npeak = levels(npeak))
  tt$xstart <- seq.int(from = 0, by = 20, length.out = length(tt$npeak))
  tt
})

table.cmps_neapk.long <- table.cmps_neapk.long %>% left_join(
  npeak_convert,
  by = "npeak"
)

table.cmps_neapk.long <- table.cmps_neapk.long %>% mutate(
  rf_status = if_else(npeak == "RF score* (Hare et al)",
    TRUE, FALSE
  )
)

table.cmps_neapk.long$Metric <-
  table.cmps_neapk.long$Metric %>%
  factor(levels = c("cmps.max", "cmps.diff", "cmps.diff_scaled", "cmps.maxbar"))

p.npeak_plot <- table.cmps_neapk.long %>%
  ggplot(aes(x = xstart + 3 * as.numeric(factor(Metric)) - 9)) +
  geom_segment(aes(
    xend = xstart + 3 * as.numeric(factor(Metric)) - 9,
    y = value, yend = 0, color = Metric
  ),
  position = position_dodge(width = 0.6)
  ) +
  geom_point(aes(y = value, color = Metric, shape = rf_status),
    alpha = 0.8,
    position = position_dodge(width = 0.5), size = 3.5
  ) +
  facet_grid(. ~ study) +
  ylab("Sums of Squares Ratio") +
  xlab("Number of Peaks") +
  coord_flip() +
  scale_x_continuous(breaks = npeak_convert$xstart, labels = npeak_convert$npeak) +
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(shape = "none") +
  scale_shape_manual(values = c(16, 1)) +
  scale_colour_manual(values = cols_2, labels = labels) +
  scale_fill_manual(values = cols_1, labels = labels)
```

>>>>

CMPS_hamby252_results_seg <- all_results_container$h252_seg
CMPS_hamby252_results_npeak <- all_results_container$h252_npeak
CMPS_hamby44_results_seg <- all_results_container$h44_seg
CMPS_hamby44_results_npeak <- all_results_container$h44_npeak
rf_results <- all_results_container$rf

@@

```{r rf-data-process, echo=FALSE}
rf_tt <- lapply(rf_results, function(rf_result) {
  with(rf_result, {
    tibble(rf.diff, rf.max, rf.maxbar) %>%
      apply(2, compute_ss_ratio, label = type_truth, MS = FALSE)
  })
}) %>%
  do.call(rbind, .) %>%
  as_tibble()

colnames(rf_tt) <- c("cmps.diff", "cmps.max", "cmps.maxbar")

rf_tt$cmps.diff_scaled <- rf_tt$cmps.diff

rf_tt$study <- c("Hamby 252", "Hamby 44")
rf_tt <- rf_tt %>% pivot_longer(
  cols = c("cmps.diff", "cmps.max", "cmps.maxbar", "cmps.diff_scaled"),
  names_to = "Metric"
)

rf_tt$npeak <- "RF score* (Hare et al)"
```

```{r ss-original-cmps, echo=FALSE}
result_container <- list()

result_container$npeak <-
  c("5* (Chen et al)", "5-3-1* (Chen et al)", "5* (Chen et al)", "5-3-1* (Chen et al)")
result_container$Metric <- c("cmps.max", "cmps.max", "cmps.maxbar", "cmps.maxbar")
result_container$study <- "Hamby 252"

dd_list <- list()

dd_list[[1]] <- tibble(
  score = c(9:13, 16:24),
  freq = c(132, 282, 115, 19, 1, 1, 3, 2, 4, 8, 12, 7, 7, 2),
  type = c(rep("KNM", 5), rep("KM", 9))
)

dd_list[[2]] <- tibble(
  score = c(1:5, 14:23),
  freq = c(7, 343, 168, 24, 7, 1, 1, 2, 3, 6, 7, 8, 8, 7, 3),
  type = c(rep("KNM", 5), rep("KM", 10))
)

dd_list[[3]] <- tibble(
  score = c(6:9, 13:19),
  freq = c(2, 210, 307, 30, 6, 3, 9, 9, 11, 7, 1),
  type = c(rep("KNM", 4), rep("KM", 7))
)

dd_list[[4]] <- tibble(
  score = c(0:2, 8:17),
  freq = c(36, 508, 5, 1, 3, 4, 2, 11, 3, 9, 8, 4, 1),
  type = c(rep("KNM", 3), rep("KM", 10))
)

result_container$dd_list <- dd_list

result_container <- result_container %>%
  as_tibble() %>%
  mutate(
    dd_table = purrr::map(dd_list, function(dd) {
      tibble(
        score = rep(dd$score, dd$freq),
        type_truth = rep(dd$type, dd$freq)
      )
    }),
    value = purrr::map_dbl(dd_table, function(dd) {
      compute_ss_ratio(dd$score, dd$type_truth, MS = FALSE)
    })
  )
```

@@

```{r plot_computation, echo=FALSE, fig.width=8, fig.height = 8, out.width='\\linewidth'}
# 252 - seg
table_252_cmps_seg <- lapply(CMPS_hamby252_results_seg$cmps.table, function(tt) {
  type_truth <- tt$type_truth
  tt %>%
    select(cmps.diff:cmps.maxbar_scaled) %>%
    apply(2, compute_ss_ratio, label = type_truth, MS = FALSE)
}) %>%
  do.call(rbind, .) %>%
  as.data.frame()

table_252_cmps_seg$seg_length <- unlist(CMPS_hamby252_results_seg$seg_length)

table_252_cmps_seg.long <- table_252_cmps_seg %>%
  pivot_longer(
    cols = c(
      cmps.diff, cmps.diff_scaled,
      cmps.max, cmps.maxbar
    ),
    names_to = "Metric"
  )

table_252_cmps_seg.long <- table_252_cmps_seg.long %>%
  mutate(study = "Hamby 252")

@@

p.seg_252_plot <- table_252_cmps_seg.long %>%
  ggplot(aes(x = seg_length + 3 * as.numeric(factor(Metric)) - 9)) +
  geom_segment(aes(
    xend = seg_length + 3 * as.numeric(factor(Metric)) - 9,
    y = value, yend = 0, colour = Metric
  )) +
  geom_point(aes(y = value, color = Metric),
    alpha = 0.8,
    position = position_dodge(width = 0.5), size = 3.5
  ) +
  scale_x_continuous(
    breaks = table_252_cmps_seg$seg_length,
    minor_breaks = NULL
  ) +
  theme_bw() +
  # theme(legend.position="none") +
  ggtitle("Results for Hamby 252") +
  ylab("Sum of Squares Ratio") +
  xlab("Segment Length") +
  scale_colour_manual(values = cols_2, labels = labels) +
  scale_fill_manual(values = cols_1, labels = labels)

# 44 - seg
table_44_cmps_seg <- lapply(CMPS_hamby44_results_seg$cmps.table, function(tt) {
  type_truth <- tt$type_truth
  tt %>%
    select(cmps.diff:cmps.maxbar_scaled) %>%
    apply(2, compute_ss_ratio, label = type_truth, MS = FALSE)
}) %>%
  do.call(rbind, .) %>%
  as.data.frame()

table_44_cmps_seg$seg_length <- unlist(CMPS_hamby44_results_seg$seg_length)

table_44_cmps_seg.long <- table_44_cmps_seg %>%
  pivot_longer(
    cols = c(
      cmps.diff, cmps.diff_scaled,
      cmps.max, cmps.maxbar
    ),
    names_to = "Metric"
  )

table_44_cmps_seg.long <- table_44_cmps_seg.long %>% mutate(
  study = "Hamby 44"
)

table.cmps_seg.long <- bind_rows(
  table_44_cmps_seg.long,
  table_252_cmps_seg.long
)

@@

# npeak - 252
table_252_cmps_npeak <- lapply(CMPS_hamby252_results_npeak$cmps.table, function(tt) {
  type_truth <- tt$type_truth
  tt %>%
    select(cmps.diff:cmps.maxbar_scaled) %>%
    apply(2, compute_ss_ratio, label = type_truth, MS = FALSE)
}) %>%
  do.call(rbind, .) %>%
  as.data.frame()

table_252_cmps_npeak$npeak <- CMPS_hamby252_results_npeak$npeaks_set %>%
  purrr::map(.f = paste, collapse = "-") %>%
  sapply(as.character)

table_252_cmps_npeak.longer <- table_252_cmps_npeak %>%
  as_tibble() %>%
  filter(npeak %in% c("5", "5-3-1", "10-8-6-4-2-1", "10-7-4-2")) %>%
  mutate(
    npeak = factor(
      npeak,
      levels = npeak[order(cmps.diff_scaled, decreasing = FALSE)]
    )
  ) %>%
  pivot_longer(
    cols = c(cmps.diff, cmps.diff_scaled, cmps.max, cmps.maxbar),
    names_to = "Metric"
  )
table_252_cmps_npeak.longer$study <- "Hamby 252"

# npeak - 44
table_44_cmps_npeak <- lapply(CMPS_hamby44_results_npeak$cmps.table, function(tt) {
  type_truth <- tt$type_truth
  tt %>%
    select(cmps.diff:cmps.maxbar_scaled) %>%
    apply(2, compute_ss_ratio, label = type_truth, MS = FALSE)
}) %>%
  do.call(rbind, .) %>%
  as.data.frame()

table_44_cmps_npeak$npeak <- CMPS_hamby44_results_npeak$npeaks_set %>%
  purrr::map(.f = paste, collapse = "-") %>%
  sapply(as.character)

table_44_cmps_npeak.longer <- table_44_cmps_npeak %>%
  as_tibble() %>%
  filter(npeak %in% c("5", "5-3-1", "10-8-6-4-2-1", "10-7-4-2")) %>%
  mutate(
    npeak = factor(
      npeak,
      levels = npeak[order(cmps.diff_scaled, decreasing = FALSE)]
    )
  ) %>%
  pivot_longer(
    cols = c(cmps.diff, cmps.diff_scaled, cmps.max, cmps.maxbar),
    names_to = "Metric"
  )
table_44_cmps_npeak.longer$study <- "Hamby 44"

@@

table_cmps_neapk.long <- bind_rows(
  table_252_cmps_npeak.longer %>% select(all_of(selected_var)),
  table_44_cmps_npeak.longer %>% select(all_of(selected_var)),
  rf_tt %>% select(all_of(selected_var)),
  result_container %>% select(all_of(selected_var))
)

order_level <- table_cmps_neapk.long %>%
  filter(Metric == "cmps.maxbar", study == "Hamby 252") %>%
  {
    .$npeak[order(.$value, decreasing = FALSE)]
  }

table_cmps_neapk.long <- table_cmps_neapk.long %>% mutate(
  npeak = factor(npeak, levels = order_level)
)

npeak_convert <- with(table_cmps_neapk.long, {
  tt <- tibble(npeak = levels(npeak))
  tt$xstart <- seq.int(from = 0, by = 20, length.out = length(tt$npeak))
  tt
})

table_cmps_neapk.long <- table_cmps_neapk.long %>% left_join(
  npeak_convert,
  by = "npeak"
)

table_cmps_neapk.long <- table_cmps_neapk.long %>% mutate(
  rf_status = if_else(npeak == "RF score* (Hare et al)",
    TRUE, FALSE
  )
)

table_cmps_neapk.long$Metric <-
  table_cmps_neapk.long$Metric %>%
  factor(levels = c("cmps.max", "cmps.diff", "cmps.diff_scaled", "cmps.maxbar"))

p.npeak_plot <- table_cmps_neapk.long %>%
  ggplot(aes(x = xstart + 3 * as.numeric(factor(Metric)) - 9)) +
  geom_segment(aes(
    xend = xstart + 3 * as.numeric(factor(Metric)) - 9,
    y = value, yend = 0, color = Metric
  ),
  position = position_dodge(width = 0.6)
  ) +
  geom_point(aes(y = value, color = Metric, shape = rf_status),
    alpha = 0.8,
    position = position_dodge(width = 0.5), size = 3.5
  ) +
  facet_grid(. ~ study) +
  ylab("Sums of Squares Ratio") +
  xlab("Number of Peaks") +
  coord_flip() +
  scale_x_continuous(breaks = npeak_convert$xstart, labels = npeak_convert$npeak) +
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(shape = "none") +
  scale_shape_manual(values = c(16, 1)) +
  scale_colour_manual(values = cols_2, labels = labels) +
  scale_fill_manual(values = cols_1, labels = labels)
```
